\textsf{Are large language models superhuman chemists?}

### 2024-03-29 src/tex/appendix.tex
include scale plot
trying to compile

@@

\begin{figure}
    \centering 
    \includegraphics{figures/llama_performance.pdf}
    \caption{\textbf{Performance of LLaMA models on the \chembench corpus.} The plot shows the fraction of correctly answered questions as a function of the model size.}
    \label{fig:llama_performance}
    \script{analyze_llama_performance.py}
\end{figure}

@@

\subsection{Leaderboard}
Our leaderboard is based on the toolchain developed for Matbench.\autocite{dunn2020benchmarking} 
Briefly, the \chembench pipeline produces standardized files in \texttt{json} format that contributors can add via pull requests to the \chembench repository.
The Markdown tables and interactive plots are automatically generated and updated on the \chembench website.

>>>>

% \begin{table}
%     \caption{\textbf{Spearman correlation coefficients for the correlation of model performance with molecular features.} The table shows the Spearman correlation}
%     \begin{tabularx}{\textwidth}{XXXX XXXX}
%         \toprule
%         topic & molecular descriptor & \(\rho\) GPT-4 & \(\rho \) Claude 3 & \( \rho \) Galactica &  \( \rho \) humans  \\ 
%         \midrule
%         number of \gls{nmr} signals & number of atoms &  \variable{output/correlation_correlation/spearman_num_atoms_gpt4_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_num_atoms_claude3_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_num_atoms_galactica_120b_is_number_nmr_peaks.txt}  & \variable{output/correlation_correlation/spearman_num_atoms_human_is_number_nmr_peaks.txt} \\
%                               & complexity & \variable{output/correlation_correlation/spearman_complexity_gpt4_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_complexity_claude3_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_complexity_galactica_120b_is_number_nmr_peaks.txt} & \variable{output/correlation_correlation/spearman_complexity_human_is_number_nmr_peaks.txt} \\
%         \midrule 
%         total electron counts & number of atoms & \variable{output/correlation_correlation/spearman_num_atoms_gpt4_is_electron_counts.txt} & \variable{output/correlation_correlation/spearman_num_atoms_claude3_is_electron_counts.txt} & \variable{output/correlation_correlation/spearman_num_atoms_galactica_120b_is_electron_counts.txt} & \variable{output/correlation_correlation/spearman_num_atoms_human_is_electron_counts.txt} \\
%         \midrule 
%         \gls{smiles}/\gls{iupac} name matching & number of atoms & \variable{output/correlation_correlation/spearman_num_atoms_gpt4_is_name.txt} & \variable{output/correlation_correlation/spearman_num_atoms_claude3_is_name.txt} & \variable{output/correlation_correlation/spearman_num_atoms_galactica_120b_is_name.txt} & \variable{output/correlation_correlation/spearman_num_atoms_human_is_name.txt} \\
%                 & complexity & \variable{output/correlation_correlation/spearman_complexity_gpt4_is_name.txt} & \variable{output/correlation_correlation/spearman_complexity_claude3_is_name.txt} & \variable{output/correlation_correlation/spearman_complexity_galactica_120b_is_name.txt} & \variable{output/correlation_correlation/spearman_complexity_human_is_name.txt} \\
%         \bottomrule
%     \end{tabularx}
%     \label{tab:correlation_coefficients}
% \end{table}

@@

\begin{figure}[htb]
    \centering 
    \includegraphics{figures/llama_performance.pdf}
    \caption{\textbf{Performance of LLaMA models on the \chembench corpus.} The plot shows the fraction of correctly answered questions as a function of the model size.}
    \label{fig:llama_performance}
    \script{analyze_llama_performance.py}
\end{figure}

@@

\subsection{Leaderboard}
Our leaderboard is based on the toolchain developed for Matbench.\autocite{Dunn_2020} 
Briefly, the \chembench pipeline produces standardized files in \texttt{json} format that contributors can add via pull requests to the \chembench repository.
The Markdown tables and interactive plots are automatically generated and updated on the \chembench website.

### 2024-03-29 src/tex/ms.tex
trying to compile

Thus, to better understand what \glspl{llm} can do for the chemical sciences and where they might be improved with further developments, evaluation frameworks are needed to allow us to measure progress and mitigate potential harms systematically.
For the development of \glspl{llm}, evaluation is currently primarily performed via standardized benchmark suites such as BigBench\autocite{srivastava2022beyond} or the LM Eval Harness.\autocite{eval-harness}
Among 204 tasks (such as linguistic puzzles), the former contains only two tasks classified as \enquote{chemistry related}, whereas the latter contains no specific chemistry tasks.
Due to the lack of widely accepted standard benchmarks, the developers of chemical language models\autocite{jablonka2024leveraging, guo2023large, ahmad2022chemberta2, Cai_2024, frey2023neural} frequently utilize language-interfaced\autocite{dinh2022lift} tabular datasets such as the ones reported in MoleculeNet,\autocite{wu2018moleculenet} Therapeutic Data Commons\autocite{huang2021therapeutics} or MatBench.\autocite{dunn2020benchmarking}
In these cases, the models are evaluated on predicting very specific properties of molecules (e.g., solubility, toxicity, melting temperature or reactivity) or on predicting the outcome of specific chemical reactions.
This, however, only gives a very limited view of the general chemical capabilities of the models.

@@

The current evaluation frameworks for chemical \glspl{llm} are primarily designed to measure the performance of the models on specific property prediction tasks. 
They cannot be used to evaluate reasoning or systems built for scientific applications. 
Thus, we had little understanding of the capabilities of \glspl{llm} in the chemical sciences.
Our work shows that carefully curated benchmarks can provide a more nuanced understanding of the capabilities of \glspl{llm} in the chemical sciences.
Importantly, our findings also illustrate that more focus is required in the development of better human-model interaction frameworks, given that models are not able to estimate their limitations.

@@

\paragraph{Parsing}
Our parsing workflow is multistep and primarily based on regular expressions.
In the case of instruction-tuned models, we first attempt to identify the \texttt{[ANSWER][\textbackslash ANSWER]} environment we prompt the model to report the answer in.
In the case of completion models, this step is skipped. From there, we attempt to extract the relevant enumeration letters (for multiple-choice questions) or numbers.
In the case of numbers, our regular expression was engineered to deal with various forms of scientific notation.
As initial tests indicated that models sometimes return integers in the form of words, e.g., \enquote{one} instead of \enquote{1}, we also implemented a word-to-number conversion using regular expressions.
If these hard-coded parsing steps fail, we use a \gls{llm}, e.g., Claude 2, to parse the completion.
%Manual verification indicates that this \gls{llm}-based parsing represent not a relevant error source and has been performed correctly in all cases we checked (randomly sampled 4 questions per topic).

>>>>

Thus, to better understand what \glspl{llm} can do for the chemical sciences and where they might be improved with further developments, evaluation frameworks are needed to allow us to measure progress and mitigate potential harms systematically.
For the development of \glspl{llm}, evaluation is currently primarily performed via standardized benchmark suites such as BigBench\autocite{srivastava2022beyond} or the LM Eval Harness.\autocite{eval-harness}
Among 204 tasks (such as linguistic puzzles), the former contains only two tasks classified as \enquote{chemistry related}, whereas the latter contains no specific chemistry tasks.
Due to the lack of widely accepted standard benchmarks, the developers of chemical language models\autocite{jablonka2024leveraging, guo2023large, ahmad2022chemberta2, Cai_2024, frey2023neural} frequently utilize language-interfaced\autocite{dinh2022lift} tabular datasets such as the ones reported in MoleculeNet,\autocite{wu2018moleculenet} Therapeutic Data Commons\autocite{huang2021therapeutics} or MatBench.\autocite{Dunn_2020}
In these cases, the models are evaluated on predicting very specific properties of molecules (e.g., solubility, toxicity, melting temperature or reactivity) or on predicting the outcome of specific chemical reactions.
This, however, only gives a very limited view of the general chemical capabilities of the models.

@@

The current evaluation frameworks for chemical \glspl{llm} are primarily designed to measure the performance of the models on specific property prediction tasks. 
They cannot be used to evaluate reasoning or systems built for scientific applications. 
Thus, we had little understanding of the capabilities of \glspl{llm} in the chemical sciences.
Our work shows that carefully curated benchmarks can provide a more nuanced understanding of the capabilities of \glspl{llm} in the chemical sciences.
Importantly, our findings also illustrate that more focus is required in developing better human-model interaction frameworks, given that models cannot estimate their limitations.

@@

\paragraph{Parsing}
Our parsing workflow is multistep and primarily based on regular expressions.
In the case of instruction-tuned models, we first attempt to identify the \texttt{[ANSWER][\textbackslash ANSWER]} environment we prompt the model to report the answer in.
In the case of completion models, this step is skipped. From there, we attempt to extract the relevant enumeration letters (for multiple-choice questions) or numbers.
In the case of numbers, our regular expression was engineered to deal with various forms of scientific notation.
As initial tests indicated that models sometimes return integers in the form of words, e.g., \enquote{one} instead of \enquote{1}, we also implemented a word-to-number conversion using regular expressions.
If these hard-coded parsing steps fail, we use a \gls{llm}, e.g., Claude 2, to parse the completion.
